---
title: "Gov 50 Final Project"
author: Greta Candreva
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Milestone 1: Project thoughts

I am interested in exploring data related to...

looking at if there is a relationship between the height, weight, and age of NBA players and the number of average career points they scored  (do shorter/taller or relatively heavier/lighter players tend to score more goals/be more successful?). 

## Milestone 2
```{r}
library(tidyverse)
library(readr)
library(stringr)


game_stats <- read_csv("all_seasons.csv", show_col_types = FALSE)
game_subset <- game_stats |>
  filter(draft_year %in% 2010:2020)
```

Does the age, weight, and/or height of NBA players affect their average number of points scored over a around a twenty year period? I hypothesize, that younger and taller players will have a greater number of average points scored, and that weight will have a large influence, if any. I think that height and age will be the most influential factors, because I think that a relatively younger and taller player will aid in a players ability to score points, therefore leading to a positive correlation. Therefore, I think that as lower ages, and taller players, will have higher average point records. My outcome variable is average points scored (pts). My explanatory variables are age (age), height (player height), and weight (player_weight). If I observe a higher record of pts for taller and younger players, this would support my hypothesis. If I observe less pts for younger and taller players, this would go against my hypothesis. Furthermore, if I observe no change in average points scored based on height and age, and/or if i observe higher pts based of weight, this would go against my hypothesis. When I run my regression of pts on the indicated variables, a positive, significant coefficient would indicate support for my hypothesis. 


#Milestone 3

```{r}
visual_1 <- game_stats |>
  ggplot(aes(
    x = age,
    y = pts
  )) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE) +
  labs(
    x = "Age in Years",
    y = "Average Points Scored"
  ) +
  theme_bw()
  
visual_1
  
```

This graph demonstrates that there appears to be a relationship between age and average points scored. I looks like average points scored peaks when a player is around 30 years old, and decreases as they get older. However, this graph does not demonstrate a very strong relationship, and this further evaluations will need to be conducted.  





